---
title: "Annotated Network Analysis Pipeline in R"
output: html_document
---

# Network Analysis Pipeline in R

This R Markdown document rebuilds the network from a CSV edge list, draws it using **fixed node positions** that match the figure, and computes:

## Node-level measures
- Raw degree
- Normalized degree centrality
- Raw betweenness centrality
- Normalized betweenness centrality
- Closeness centrality
- Eigenvector centrality
- Clustering coefficient

## Graph-level measures
- Average clustering coefficient
- Global clustering coefficient
- Diameter
- Average path length
- Density

At the end, the pipeline saves:
- `network_centrality_results.csv` for node-level results
- `network_global_metrics.csv` for graph-level results

# 1. Install and load packages

We use:

- `readr` to read CSV files
- `dplyr` to build clean result tables
- `igraph` to construct, analyze, and plot the network

```{r setup, message=FALSE, warning=FALSE}
packages <- c("readr", "dplyr", "igraph")

install.packages(packages)

library(readr)
library(dplyr)
library(igraph)
```

# 2. Read the CSV file

The CSV is expected to have this structure:

```csv
source,target
A,B
A,C
...
```

Each row represents one **undirected edge**.

```{r read-csv}
# Change this filename if needed
csv_filename <- "network_edges.csv"

edges <- read_csv(csv_filename, show_col_types = FALSE)
edges
```

# 3. Construct the network

We build an **undirected graph** because the original figure does not show arrow directions.

```{r build-graph}
g <- graph_from_data_frame(edges, directed = FALSE)

cat("Number of nodes:", vcount(g), "\n")
cat("Number of edges:", ecount(g), "\n")
cat("Nodes:", paste(sort(V(g)$name), collapse = ", "), "\n")
```

# 4. Define fixed node positions

Instead of using an automatic layout, we manually define coordinates so the plotted graph resembles the source image more closely.

```{r fixed-positions}
node_order <- c("A", "B", "C", "D", "E", "F", "G", "H")

fixed_positions <- data.frame(
  name = node_order,
  x = c(-2.0,  0.0, -1.0, -2.5, -1.2,  0.2,  1.5,  1.5),
  y = c( 1.0,  1.2,  0.0, -0.5, -1.5, -0.8, -1.5,  0.0)
)

fixed_positions
```

# 5. Draw the network

The layout matrix must follow the exact order of the graph vertices.

```{r plot-network, fig.width=8, fig.height=6}
# Reorder fixed coordinates so they match the graph's internal vertex order
layout_df <- fixed_positions[match(V(g)$name, fixed_positions$name), ]
layout_matrix <- as.matrix(layout_df[, c("x", "y")])

plot(
  g,
  layout = layout_matrix,
  vertex.size = 35,
  vertex.color = "lightgray",
  vertex.frame.color = "black",
  vertex.label.color = "black",
  vertex.label.cex = 1.2,
  edge.color = "black",
  main = "Network Graph (Fixed Layout)"
)
```

# 6. Compute node-level measures

Below is a short interpretation of each measure:

- **Raw degree**: number of direct neighbors
- **Normalized degree centrality**: raw degree divided by the maximum possible degree
- **Raw betweenness**: how often a node lies on shortest paths between other nodes, unnormalized
- **Normalized betweenness**: scaled betweenness
- **Closeness centrality**: how near a node is to all others
- **Eigenvector centrality**: gives more weight to connections to important nodes
- **Clustering coefficient**: how connected a node's neighbors are to one another

```{r node-metrics}
# Raw degree
degree_raw <- degree(g, mode = "all", normalized = FALSE)

# Normalized degree
degree_norm <- degree(g, mode = "all", normalized = TRUE)

# Raw betweenness
betweenness_raw <- betweenness(g, directed = FALSE, normalized = FALSE)

# Normalized betweenness
betweenness_norm <- betweenness(g, directed = FALSE, normalized = TRUE)

# Closeness centrality
closeness_vals <- closeness(g, normalized = TRUE)

# Eigenvector centrality
eigenvector_vals <- eigen_centrality(g, directed = FALSE, scale = TRUE)$vector

# Local clustering coefficient
clustering_local <- transitivity(g, type = "localundirected")
```

# 7. Combine node-level results into one table

```{r node-results}
node_results <- tibble(
  Node = V(g)$name,
  Degree_raw = as.numeric(degree_raw),
  Degree_norm = as.numeric(degree_norm),
  Betweenness_raw = as.numeric(betweenness_raw),
  Betweenness_norm = as.numeric(betweenness_norm),
  Closeness = as.numeric(closeness_vals),
  Eigenvector = as.numeric(eigenvector_vals),
  ClusteringCoeff = as.numeric(clustering_local)
) %>%
  arrange(desc(Degree_raw), desc(Betweenness_raw)) %>%
  mutate(
    Degree_norm = round(Degree_norm, 6),
    Betweenness_raw = round(Betweenness_raw, 6),
    Betweenness_norm = round(Betweenness_norm, 6),
    Closeness = round(Closeness, 6),
    Eigenvector = round(Eigenvector, 6),
    ClusteringCoeff = round(ClusteringCoeff, 6)
  )

node_results
```

# 8. Save node-level results to CSV

```{r save-node-results}
write_csv(node_results, "network_centrality_results.csv")
```

# 9. Compute graph-level measures

These metrics describe the **whole network**, not individual nodes.

- **Average clustering coefficient**: average of local clustering across all nodes
- **Global clustering coefficient**: triangle density over all connected triples
- **Diameter**: the longest shortest-path distance in the network
- **Average path length**: average shortest-path distance between all node pairs
- **Density**: proportion of realized edges out of all possible edges

```{r global-metrics}
avg_clustering <- transitivity(g, type = "average")
global_clustering <- transitivity(g, type = "globalundirected")
diameter_val <- diameter(g, directed = FALSE, weights = NA)
avg_path_length <- mean_distance(g, directed = FALSE)
density_val <- edge_density(g, loops = FALSE)

global_results <- tibble(
  Metric = c(
    "AverageClusteringCoeff",
    "GlobalClusteringCoeff",
    "Diameter",
    "AveragePathLength",
    "Density"
  ),
  Value = c(
    avg_clustering,
    global_clustering,
    diameter_val,
    avg_path_length,
    density_val
  )
) %>%
  mutate(Value = round(Value, 6))

global_results
```

# 10. Save graph-level results to CSV

```{r save-global-results}
write_csv(global_results, "network_global_metrics.csv")

```


